PhD Title \\ PhD Thesis

### 2024-07-24 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. Because 
of thermodynamics, the two can only remain in equilibrium while 
below a specific concentration ($C_{eq}$) - below which 
the chemical potential $\mu$ for a miscible solution 
is greater than the potential required to separate 
the two phases. Once $C_{eq}$ is exceeded there is a chemical
 potential difference driving the solution to separate the 
two phases. Since different combinations of solute and solvent
will have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by using 
'supersaturation', one form of measuring supersaturation is shown 
below \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}
Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
is the equilibrium concentration at temperature $T$.
While the solution remains supersaturated there is a 
chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, 
the first formation of the crystal is referred to as the 
nucleus and understanding its formation of has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation
of the system.

\subsection{Primary vs Secondary nucleation}
From an industrial perspective, the nucleation process can
be broadly categorised into either primary of secondary 
nucleation. The former being described as the formation of an
initial nucleus within the bulk phase; absent of any external
stimuli primary nucleation is considered stochastic and is typically
undesirable for industrial applications where short and consistent 
residence times are desirable. At a small scale one can estimate the 
nucleation rate by making repeated measurements of sample solutions 
and seeing how many have nucleated after a given time, giving us a 
Poisson distribution.

@@

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
The newly formed phase has a lower chemical potential than the 
surrounding solution, reducing the free energy of the system; 
at the same time, the formation of a new interface is resited 
by the bulk phase due to surface tension. The net free energy 
of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
\end{align}
Where $v$ is the approximate volume of an individual molecule, $k_B$
is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
tension of the bulk solution. Plotting the free energy of the system
against nucleus size reveals a critical size above which the gain 
in free energy exceeds the interfacial tension. 
\begin{figure}[h!]
	\label{fig:free_energy}
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
\end{figure}

The maximum value of $\Delta G_{tot}$ is the free energy barrier 
that any newly formed nucleus needs to overcome in order to stabilise.
The nucleation rate (the volume of new crystalline material formed per
unit time), is therefore commonly defined as being dependent on the 
energy barrier $\Delta G^*$:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}
Where $A$ is a pre-factor that can be fine tuned to the exact demands
of the system (typically involving Zeldovich Factor, $Z$) the free 
energy barrier can be found by finding the turning point of $\Delta G_{tot}$.

So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy. The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius. This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}

Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows. Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.

There are several results that indicate that the CNT is not a good fit for nucleation rates:
\textcolor{red}{
	\begin{itemize}
		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 \cite. 
		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass. We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure. To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ]. 
		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled. Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases. [ ]
	\end{itemize}
}

CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate. However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case. Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, colloid simulations showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) interactions allow for the formation of a liquid-liquid metastable phase from which a new solid phase could form \cite{Anderson2002, Karthika2016}. \textit{in situ} techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.  It can be understood by Oswald's rule \cite{Ostwald1897}, which says that any crystallising system does not 
take the path to the lowest possible energy state but instead first transitions to the state with the smallest free energy barrier, further transitions can still occur but this method minimises the overall free energy cost. If we consider the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider a case where the supersaturation is low ($S\approx1.01$) then a dense liquid droplet will have a lower chemical potential than the surrounding fluid

\section{Crystallisation methods}
\subsection{Cooling Crystallisation}
For some binary mixtures the supersaturation is heavily dependent on 
the solution temperature, therefore a simple method of producing crystals 
is by cooling the mixture to induce crystal formation. At ambient 
temperatures the solute concentration is too high to be fully incorporated
into the solution ($S\gg 1$), after heating however the solute is fully dissolved
into the solution ($S<1$). Now as the mixture is allowed to cool to room
temperature the supersaturation will increase until crystal formation begins,
the rate of cooling drastically influencing the size and number of crystals
produced.

@@

\begin{figure}[h!]
	\label{fig:MSZW}
	\includegraphics[width=\linewidth]{MSZW.png}
	\caption{Typical Concentration vs Temperature plot with the solubility curve (black) and metastable zone (red). 3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
\end{figure}

@@

\section{Optical Tweezers}
\subsection{Background}
Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'. The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis. Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.

@@

\subsection{Optical Rotation}
For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum. Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components. The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}. However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}. Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}. This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.

For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM. Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus. This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam. Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM. However, by far the best method for inducing rotation is using birefringent materials.

Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation. For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}. Calculating the optical torque applied to a birefringent material is given via:

@@

Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.  The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation. The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle. Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion. Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum.

\subsection{Shear induced Nucleation}
It has long been known that solution nucleation is influenced by shearing 
of the fluid, from stirring agents in the container (i.e. propellers) to 
the container boundary layers, however there has yet to be a clear mechanism 
by which said shearing effects the nucleation event. Theoretical research 
into shear induced nucleation suggests that there should be a slight increase 
in the nucleation rate at low shear rates, reaching a maximum increase in 
nucleation rate, and then at higher shear rates the nucleation rate begins 
to drop off. This has been shown theoretically for both simple colloidal
\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
\cite{Goswami2020}; however, no experimental work into these systems has 
been conducted to prove this is the case.  There is some experimental 
evidence for this phenomena in simple salt and protein solutions - though 
the authors emphasise that mechanical agitation cannot be ruled out - 
there has not been a exhaustive study into the shearing effects apart from 
in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
the highest nucleation rate. Using the theoretical model established in 
\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
a nucleus undergoing shearing, accounting for the fact that a nucleus' 
growth is undergoing competition between flow-mediated molecular transport 
and the strain applied by the flow field which inhibits the growth of the 
nucleus. There central conclusion (from both the theoretical and experimental 
results) is that there is an optimal shear rate in which the nucleation 
rate is maximised. However, a question that arises from this result, if 
there is a optimal shear rate in which molecular transport is maximised 
and strain is minimised, then surely there should also be a shear rate 
in which the molecular transport and strain are equal - allowing one to 
suspend a nucleus at a constant radius. In this scenario, the molecular 
transport would prevent the nucleus from dissolving, but the strain would 
prevent the nucleus from growing. This however would require one to be 
able to apply a continuous shear rate to a targeted nucleus with high 
precision, there is also no model for an individual nucleus undergoing growth.

\section{Aims and Objectives}
Overall the aim of the PhD is to study viability of using micro-rotors to
generate localised fluid flow around the beam focus. The results are reported
in chapter 3, this is then succeeded by experimental work where we 
instead using a galvano-mirror to generate a moving beam focus within an 
undersaturated solution to control the crystal growth front. In addition,
computer simulations into asymmetric dimers and their applications as complex
micro-rotors are discussed in chapter 4. Lastly two novel methods of measuring
rotational motion are discussed in chapter 5; firstly via a novel detection
fibre method that allows for instantaneous measurements of the orientational
behaviour of optically trapped ellipsoids/dimers; and secondly we create 
a simulative quadrant photo diode that replicates laboratory results, utilising 
linear regression techniques we measure the change in orientation in order to
measure the optical torque applied to a non-birefringent particle.

>>>>

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. Because 
of thermodynamics, the two can only remain in equilibrium while 
below a specific concentration ($C_{eq}$) - below which 
the chemical potential $\mu$ for a miscible solution 
is greater than the potential required to separate 
the two phases. Once $C_{eq}$ is exceeded there is a chemical
 potential difference driving the solution to separate the 
two phases. Since different combinations of solute and solvent
will have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by using 
'supersaturation', one form of measuring supersaturation is shown 
below \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
is the equilibrium concentration at temperature $T$.
While the solution remains supersaturated there is a 
chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, 
the first formation of the crystal is referred to as the 
nucleus and understanding its formation of has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation
of the system.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can
be broadly categorised into either primary of secondary 
nucleation. The former being described as the formation of an
initial nucleus within the bulk phase; absent of any external
stimuli primary nucleation is considered stochastic and is typically
undesirable for industrial applications where short and consistent 
residence times are desirable. At a small scale one can estimate the 
nucleation rate by making repeated measurements of sample solutions 
and seeing how many have nucleated after a given time, giving us a 
Poisson distribution.

@@

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
The newly formed phase has a lower chemical potential than the 
surrounding solution, reducing the free energy of the system; 
at the same time, the formation of a new interface is resited 
by the bulk phase due to surface tension. The net free energy 
of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
\end{align}

Where $v$ is the approximate volume of an individual molecule, $k_B$
is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
tension of the bulk solution. Plotting the free energy of the system
against nucleus size reveals a critical size above which the gain 
in free energy exceeds the interfacial tension. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
	     \label{fig:free_energy}
\end{figure}

The maximum value of $\Delta G_{tot}$ is the free energy barrier 
that any newly formed nucleus needs to overcome in order to stabilise.
The nucleation rate (the volume of new crystalline material formed per
unit time), is therefore commonly defined as being dependent on the 
energy barrier $\Delta G^*$:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}

Where $A$ is a pre-factor that can be fine tuned to the exact demands
of the system (typically involving Zeldovich Factor, $Z$) the free 
energy barrier can be found by finding the turning point of $\Delta G_{tot}$.

CNT is often regarded as a good description of the macro system, its obvious
that for all crystallization systems there is an inherent energy barrier that 
dictates the nucleation rate. Where it falters is in its predictive ability, 
both in estimating nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in 
the structure of newly formed nuclei \cite{Lee1999, Yau2001}. Recent studies 
suggest classical nucleation is merely one of many possible pathways that can 
be taken to produce a structured crystalline phase. Prompting the development of alternative theories to better describe the nucleation process.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, colloid simulations 
showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) 
interactions allow for the formation of a liquid-liquid metastable phase from 
which a new solid phase could form \cite{Anderson2002, Karthika2016}. 
\textit{in situ} techniques for studying nucleation several papers reported 
the presence of stable liquid-like clusters that formed prior to nucleation
\cite{Savage2009, Wolde1997, Soga1999}. It can be understood by Oswald's rule
\cite{Ostwald1897}, which says that any crystallising system does not immediately 
take the path to the lowest possible energy state but instead first transitions 
to the state with the smallest free energy barrier, further transitions can still 
occur but this pathway minimises the overall free energy cost. If we consider 
the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider 
a case where the supersaturation is low ($S\approx1.01$) then a dense liquid 
droplet will have a lower chemical potential than the surrounding fluid but 
the crystalline phase will have a subsequently lower chemical potential than 
the dense liquid phase. Initially it was suspected that the formation of the 
dense liquid phase was a result of stochastic density fluctuations in the system; 
further simulations suggest that the the local bond order is a stronger driving 
force than the local density \cite{Tan2013}.

The current research into two step nucleation (or now more commonly referred 
to as multi-step nucleation theory) is developing a robust framework to 
describe what nucleation pathway will occur given the initial conditions. 
Reviews of all currently documented nucleation pathways identified highlighted 
the fact the need for a the development of \textit{in situ} techniques that 
can induce nucleation locally but can also reliably identify the nucleation 
pathway across a broad range of experimental conditions \cite{Karthika2016, Fu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Crystallisation methods}
\subsection{Cooling Crystallisation}
For some binary mixtures the supersaturation is heavily dependent on 
the solution temperature, therefore a simple method of producing crystals 
is by cooling the mixture to induce crystal formation. At ambient 
temperatures the solute concentration is too high to be fully incorporated
into the solution ($S\gg 1$), after heating however the solute is fully dissolved
into the solution ($S<1$). Now as the mixture is allowed to cool to room
temperature the supersaturation will increase until crystal formation begins,
the rate of cooling drastically influencing the size and number of crystals
produced.

@@

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{MSZW.png}
	\caption{Typical Concentration vs Temperature plot with the solubility curve (black) and metastable zone (red). 3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
	\label{fig:MSZW}
\end{figure}

@@

\subsection{Evaporative Crystallisation}
In situations where control of the final product size or shape is not the main
focus, often the cheapest method of producing a crystalline product is simply
to allow the solvent evaporate and separate from the solute. Depending on the 
total volume of solvent to evaporate this process can take on time scale of 
several days to complete.

\section{Optical Tweezers}
\subsection{Background}
Optical tweezing has been a field of applied optics ever since the 1970s
when Ashkin \cite{Ashkin1970} first showed that focused light was capable 
of trapping micron sized particles due to light exerting 'radiation pressure'. 
The working principle was that a light source such as a laser could trap 
small objects within a 2D plane, as long as the light source had an 
approximately Gaussian profile. Soon after, Ashkin showed that the introduction 
of a microscope objective would allow one to focus the light source to a 
diffraction limited point that would stably trap small objects within a 
confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study 
biological material and would later be used to probe microscopic properties 
such as the formation of colloidal aggregates \cite{Yi2021} to the drag 
forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the 
predictable behaviour of light, optical tweezers have become essential for 
measuring and exerting precise forces on the magnitude of pico-newtons 
allowing one to probe the material properties of the smallest materials.

@@

\subsection{Optical Rotation}
For any electromagnetic field it is possible to transfer both linear and 
angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
is said to have both orbital and spin momentum. Though there is some 
debate on how to decompose the total momentum into these two components 
\cite{Bruce2020, Svak2018}, for this thesis we do not need to calculate 
the exact quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the wavefront 
of the particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation pressure 
that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
non-uniform wave fronts meaning the orbital momentum has both angular and
linear components; depending on the relative size of the target particle 
one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.

Spin angular momentum (SAM) is attributed to the spin density of the field, 
early research has shown that the spin density is non-zero for any beam despite
the fact that the total SAM transferred to a medium is 0 \cite{Svak2018, 
Bliokh2014}. This has sparked debate if SAM is even a physical quantity as 
it does not aid in the transport of energy directly \cite{Bliokh2014} and 
so cannot be directly observed in some cases despite being non-zero. This 
paradox is resolved by representing the wave as an array of spin momentum 
loops that all together cancel one-another out when the medium is homogeneous.
Spacial inhomogeneities cause these spin loops to non longer be equal, resulting
in non-zero spin density, anisotropic mediums (such as birefringent crystal 
lattices) experience a transfer of spin angular momentum allowing for optical
torque transfer. It has even been shown that in the case of homogeneous spheres, 
the spin component of a focused beam can result in additional orbital angular
momentum via spin-to-orbital conversion \cite{Yevick2017}.

Birefringence is a material property often seen in crystalline materials, 
if the crystal lattice has a different structures when viewed at different 
orientations then light will be refracted differently depending on its 
polarisation. For circularly polarised light the inhomogeneity results in 
a high degree of SAM being transferred to the target object \cite{Parkin2009,
Arita2016}, this has been exploited to rotate microspheres as fast as 1000 Hz 
while suspended in a bulk medium \cite{Arita2016} as well as a means of 
measuring the local temperature and shear response of said medium 
\cite{Millen2014, RodriguezSevilla2018}. Calculating the optical torque 
applied to a birefringent material is given via:

@@

Where $\theta$ is the angle between the particle's long axis and the Poynting
vector of the local EM field, and $\phi$ is the phase shift in the EM field.  
The first term represents the 'orientational' torque which is due to the target 
particle being aligned along the EM field, when aligned $\theta=0$ meaning 
the entire term is negligible for particle's with a stable orientation. The 
second term is due purely to the polarisation of the optical trap, for 
circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred 
to the target particle. Birefringence can also be induced if the target 
particle has an anisotropic shape, in particular if the particle shape is 
elongated along one major axis; the go to particle shape is a silica dimer 
(two spheres tangentially attached) due to silica's stability and strong 
adhesion. Using a silica dimer research groups have achieved a rotation 
frequency in the realm of several GHz \cite{Ahn2018} in a vacuum.

\subsection{Shear induced Nucleation}
It has long been known that fluid shear rate plays a role in influencing
nucleation; however, the exact relationship between shear rate and nucleation
rate has only been recently understood for specific solutions. Theoretical research 
into shear induced nucleation suggests that there should be a slight increase 
in the nucleation rate at low shear rates, reaching a maximum increase in 
nucleation rate, and then at higher shear rates the nucleation rate begins 
to drop off. This has been shown theoretically for both simple colloidal
\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
\cite{Goswami2020}; however, no experimental work into these systems has 
been conducted to prove this is the case.  There is some experimental 
evidence for this phenomena in simple salt and protein solutions - though 
the authors emphasise that mechanical agitation cannot be ruled out - 
there has not been a exhaustive study into the shearing effects apart from 
in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
the highest nucleation rate. Using the theoretical model established in 
\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
a nucleus undergoing shearing, accounting for the fact that a nucleus' 
growth is undergoing competition between flow-mediated molecular transport 
and the strain applied by the flow field which inhibits the growth of the 
nucleus. There central conclusion (from both the theoretical and experimental 
results) is that there is an optimal shear rate in which the nucleation 
rate is maximised. However, a question that arises from this result, if 
there is a optimal shear rate in which molecular transport is maximised 
and strain is minimised, then surely there should also be a shear rate 
in which the molecular transport and strain are equal - allowing one to 
suspend a nucleus at a constant radius. In this scenario, the molecular 
transport would prevent the nucleus from dissolving, but the strain would 
prevent the nucleus from growing. This however would require one to be 
able to apply a continuous shear rate to a targeted nucleus with high 
precision, there is also no model for an individual nucleus undergoing growth.

\section{Aims and Objectives}
Overall the aim of the PhD is to study viability of using micro-rotors to
generate localised fluid flow around the beam focus. The results are reported
in chapter 3, this is then succeeded by experimental work where we 
instead using a galvano-mirror to generate shear flow. While overall unsuccessful
the addition of a moving beam focus showed that the growth of a nucleus can be localised around the trap focus.

In addition, computer simulations studying the dynamics of asymmetric dimers 
and their applications as complex micro-rotors are discussed in chapter 4. 
Lastly two novel methods of measuring rotational motion are discussed in 
chapter 5; firstly via a novel detection fibre method that allows for 
instantaneous measurements of the orientational behaviour of optically 
trapped ellipsoids/dimers; and secondly we create a simulative quadrant 
photo diode that replicates laboratory results, utilising linear regression 
techniques we measure the change in orientation in order to measure the 
optical torque applied to a non-birefringent particle.
